Preprocessing ADNI Structural MRI Data for LPME

### 2024-06-14 code/03_adni_mri_preprocess.qmd
use here package for managing working directory in rmarkdown file

```{r, message=FALSE}
library(foreach)      # used for parallelization
library(fslr)         # includes functions for registration and segmentation
library(oro.dicom)    # dicom file operations
library(oro.nifti)    # nifti file operations
library(neurobase)    # added functions for nifti operations
library(stringr)      # string and regex operations
library(tidyverse)    # used for dataframe construction
```

@@

parallel::stopCluster(cl = cl) # stop parallelized cluster
```

@@

```{r}
patnos <- list.dirs("data/adni_processed_fsl", recursive = FALSE, full.names = FALSE)
processed_dirs <- list.dirs("data/adni_processed_fsl", recursive = FALSE)

@@

```{r, eval=FALSE}
for (dir_idx in 1:length(processed_dirs)) {
  # for each subject

@@

      # repeat for right thalamus, index 9
      temp_rthal <- temp_segs[, , , 9]
      temp_rthal_vol <- seg_vols[9] %>%
        str_replace(
          pattern = "([^\\s]+)",
          replacement = ""
        ) %>%
        gsub(
          pattern = " ",
          replacement = ""
        ) %>%
        as.numeric()
      rthal_vol <- c(rthal_vol, temp_rthal_vol)
      rthal_idx <- which(temp_rthal > 0, arr.ind = TRUE)
      surface <- rep(FALSE, nrow(rthal_idx))
      for (dim in 1:3) {
        for (dim2 in 1:3) {
          if (dim != dim2) {
            dim3 <- seq(1, 3, 1)[!(1:3 %in% c(dim, dim2))]
            unique_vals <- unique(rthal_idx[, c(dim, dim2)])
            for (i in 1:nrow(unique_vals)) {
              vals <- rthal_idx[rthal_idx[, dim] == unique_vals[i, 1] & rthal_idx[, dim2] == unique_vals[i, 2], dim3]
              min_vals <- vector()
              max_vals <- vector()
              for (v in vals) {
                if (!((v-1) %in% vals)) {
                  min_vals <- c(min_vals, v)
                } else if (!((v+1) %in% vals)) {
                  max_vals <- c(max_vals, v)
                }
              }
              surface_vals <- c(min_vals, max_vals)
              for (v in surface_vals) {
                surface[which(rthal_idx[, dim] == unique_vals[i, 1] & rthal_idx[, dim2] == unique_vals[i, 2] & rthal_idx[, dim3] == v, arr.ind = TRUE)] <- TRUE
              }
            }
          } else {
            next
          }
        }
      }
      rthal_temp <- cbind(
        patnos[dir_idx],
        scan_dates[scan_idx],
        rthal_idx %*% diag(pixdim(as.nifti(temp_rthal))[2:4]),
        temp_rthal[rthal_idx]
      )
      rthal_surface_temp <- cbind(
        patnos[dir_idx],
        scan_dates[scan_idx],
        rthal_idx[surface, ] %*% diag(pixdim(as.nifti(temp_rthal))[2:4]),
        temp_rthal[rthal_idx[surface, ]]
      )
      rthal <- rbind(rthal, rthal_temp)
      rthal_surface <- rbind(rthal_surface, rthal_surface_temp)
    }
  }
}
```

@@

write_csv(hipp_info, "data/adni_fsl_hipp_info.csv")
write_csv(thal_info, "data/adni_fsl_thal_info.csv")

write_csv(lhipp, "data/adni_fsl_lhipp.csv")
write_csv(rhipp, "data/adni_fsl_rhipp.csv")

write_csv(lthal, "data/adni_fsl_lthal.csv")
write_csv(rthal, "data/adni_fsl_rthal.csv")

write_csv(lhipp_surface, "data/adni_fsl_lhipp_surface.csv")
write_csv(rhipp_surface, "data/adni_fsl_rhipp_surface.csv")

write_csv(lthal_surface, "data/adni_fsl_lthal_surface.csv")
write_csv(rthal_surface, "data/adni_fsl_rthal_surface.csv")
```

>>>>

```{r, message=FALSE}
library(foreach)      # used for parallelization
library(fslr)         # includes functions for registration and segmentation
library(here)
library(oro.dicom)    # dicom file operations
library(oro.nifti)    # nifti file operations
library(neurobase)    # added functions for nifti operations
library(stringr)      # string and regex operations
library(tidyverse)    # used for dataframe construction
```

@@

parallel::stopCluster(cl = cl) # stop parallelized cluster

```

@@

```{r}
patnos <- list.dirs(here("data/adni_processed_fsl"), recursive = FALSE, full.names = FALSE)
processed_dirs <- list.dirs(here("data/adni_processed_fsl"), recursive = FALSE)

@@

```{r, eval=FALSE}

for (dir_idx in 1:length(processed_dirs)) {
  # for each subject

@@

      # repeat for right thalamus, index 9
      temp_rthal <- temp_segs[, , , 9]
      temp_rthal_vol <- seg_vols[9] %>%
        str_replace(
          pattern = "([^\\s]+)",
          replacement = ""
        ) %>%
        gsub(
          pattern = " ",
          replacement = ""
        ) %>%
        as.numeric()
      rthal_vol <- c(rthal_vol, temp_rthal_vol)
      rthal_idx <- which(temp_rthal > 0, arr.ind = TRUE)
      surface <- rep(FALSE, nrow(rthal_idx))
      for (dim in 1:3) {
        for (dim2 in 1:3) {
          if (dim != dim2) {
            dim3 <- seq(1, 3, 1)[!(1:3 %in% c(dim, dim2))]
            unique_vals <- unique(rthal_idx[, c(dim, dim2)])
            for (i in 1:nrow(unique_vals)) {
              vals <- rthal_idx[rthal_idx[, dim] == unique_vals[i, 1] & rthal_idx[, dim2] == unique_vals[i, 2], dim3]
              min_vals <- vector()
              max_vals <- vector()
              for (v in vals) {
                if (!((v-1) %in% vals)) {
                  min_vals <- c(min_vals, v)
                } else if (!((v+1) %in% vals)) {
                  max_vals <- c(max_vals, v)
                }
              }
              surface_vals <- c(min_vals, max_vals)
              for (v in surface_vals) {
                surface[which(rthal_idx[, dim] == unique_vals[i, 1] & rthal_idx[, dim2] == unique_vals[i, 2] & rthal_idx[, dim3] == v, arr.ind = TRUE)] <- TRUE
              }
            }
          } else {
            next
          }
        }
      }
      rthal_temp <- cbind(
        patnos[dir_idx],
        scan_dates[scan_idx],
        rthal_idx %*% diag(pixdim(as.nifti(temp_rthal))[2:4]),
        temp_rthal[rthal_idx]
      )
      rthal_surface_temp <- cbind(
        patnos[dir_idx],
        scan_dates[scan_idx],
        rthal_idx[surface, ] %*% diag(pixdim(as.nifti(temp_rthal))[2:4]),
        temp_rthal[rthal_idx[surface, ]]
      )
      rthal <- rbind(rthal, rthal_temp)
      rthal_surface <- rbind(rthal_surface, rthal_surface_temp)
    }
  }
}

```

@@

write_csv(hipp_info, here("data/adni_fsl_hipp_info.csv"))
write_csv(thal_info, here("data/adni_fsl_thal_info.csv"))

write_csv(lhipp, here("data/adni_fsl_lhipp.csv"))
write_csv(rhipp, here("data/adni_fsl_rhipp.csv"))

write_csv(lthal, here("data/adni_fsl_lthal.csv"))
write_csv(rthal, here("data/adni_fsl_rthal.csv"))

write_csv(lhipp_surface, here("data/adni_fsl_lhipp_surface.csv"))
write_csv(rhipp_surface, here("data/adni_fsl_rhipp_surface.csv"))

write_csv(lthal_surface, here("data/adni_fsl_lthal_surface.csv"))
write_csv(rthal_surface, here("data/adni_fsl_rthal_surface.csv"))
```
